---
title: "Introduction to Quarto"
author: "Lauri"
date: "2026-01-30"
format: html
---

# Introduction

Welcome to this introductory **Quarto** document, designed to demonstrate the basics of Markdown formatting and R code chunks. In this document, you'll learn how to structure your content and embed code for performing ecological data analysis in R.

## Getting Started with Markdown

Markdown allows you to format your text in various ways:

-   Use `#` for headings, as shown above.
-   **Bold text** and *italicize text*.
-   Bullet points are created with `-` or `*`.
-   Numbered lists like this:
    1.  Item one
    2.  Item two

You can also add links, such as [Quarto documentation](https://quarto.org/docs/), to help students learn more.

## Importing Libraries and Data

Below, we use R to load commonly used libraries and also load some example data.

```{r}
#| label: Load-necessary-libraries
#| echo: false
#| output: false
library(tidyverse)
```

Now let's load a sample dataset. For this example, we'll use the built-in iris dataset to illustrate some basic data manipulation techniques.

```{r}
#| label: Load-data
data_test <- 
  iris %>% 
  dplyr::glimpse()
```

## Basic Data Manipulation

This example demonstrates a simple filtering operation in R. Let's select only observations of the species setosa.

```{r}
#| label: Filter-data
data_setosa <- 
  data_test %>% 
  dplyr::filter(Species == "setosa")

dplyr::glimpse(data_setosa)
```

## Data Visualization

Visualizations are essential for ecological data. Below, we create a scatter plot to examine the relationship between Sepal.Length and Sepal.Width for all species in the dataset.

```{r}
#| label: Scatter-plot
# Scatter plot
data_setosa  %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = Sepal.Length,
       y = Sepal.Width)) +
  ggplot2::geom_point() +
  ggplot2::labs(
    title = "Sepal Length vs. Sepal Width",
       x = "Sepal Length",
       y = "Sepal Width") +
  ggplot2::theme_minimal()
```
